Time-series modeling of epidemics in complex populations: detecting changes in incidence volatility over time

### 2025-02-13 ms-3/ms.tex
Complete figure 3 and writing

\clearpage
\begin{figure}[!ht]
\includegraphics[width=1\textwidth]{fig3}
\caption{caption.}\label{fig3}
\end{figure}
\clearpage

>>>>

\clearpage
\begin{figure}[!ht]
\includegraphics[width=1\textwidth]{fig3}
\caption{Incidence and dispersion between Jan 4, 2020 and March 18, 2023, in large counties in the US.
a: Mean COVID-19 cases of the 144 US counties over time. 
b: Mean $\log_{10}\theta$ of the 144 US counties over time.  
c: $\log_{10}(cases)$ over time for each of the 144 counties.
d: $\log_{10}\theta$ over time for each of the 144 counties.
e: Expected value of $\log_{10}\theta_0$ under the null model, assuming a reporting rate of 0.5.
f: LRT p-values over time for each location.}\label{fig3}
\end{figure}
\clearpage

Our method forms part of a larger interest in investigating variability in infections as an important attribute of epidemic time series using novel metrics.
For instance, burst-tree decomposition of time series has also facilitated computation of a burst-size distribution for a series given a specified time window~\citep{jo_burst-tree_2020}, allowing comparison of variability within one location over time. 
Spatial variation in superspreading potential has been investigated through risk maps of superspreading environments~\citep{loo_identification_2021}, and future work could investigate the correspondence between dispersion in case count time series, as quantified here, and indicators of a high risk of superspreading, with the potential to further illucidate drivers of transmission risk across scales, and more finely resolve landscapes of susceptibility.
Additionally, as population-wide disease control may be less effective than those which are focussed to individuals in high-transmission contexts~\citep{lloyd2005superspreading}, identifying candidate time periods when transmission heterogeneity is high may catalyze the development of more effective control strategies, particularly those that connect vulnerable populations with resources at critical times.
The finding that dispersion increased rather than decreased during the 2022 surge challenges theoretical expectations and suggests that fundamental assumptions about the scaling of transmission dynamics may require reevaluation. 
One hypothesis is that transmission heterogeneity could play a role in driving large surges, amplifying incidence beyond what homogeneous models predict. 
Future work could investigate whether bursts of highly clustered transmission events generate feedback that accelerates epidemic spread, which, if true, could refine predictive models of contagion dynamics in complex populations.
